---
title: "Class 11: Structural Bioinformatics"
author: "Stephen Curry"
date: "11/5/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The PDB database for biomolecular structure data
Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” >
“by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project
and determine the percentage of structures solved by X-Ray and Electron Microscopy. Also can
you determine what proportion of structures are protein? 

Download CSV
```{r}
data <- read.csv("Data Export Summary.csv")
data
```

```{r}
(data$Total)
```


```{r}
sum(data$Total)
```

```{r}
sum(data$Proteins)
```

```{r}
round(sum(data$Proteins)/sum(data$Total) * 100,2)
```

```{r}
(data$Total/sum(data$Total)) * 100
```

Q2:3039

```{r}
library(bio3d)
```

```{r}
pdb <- read.pdb("1hsg.pdb")
pdb
```

```{r}
a.inds <- atom.select(pdb, chain = "A")
a.inds
```


```{r}
ligand <- atom.select(pdb, "protein", value = TRUE)
write.pdb(ligand, file = "1hsg_ligand.pdb")
```



