---
title: "Class9"
author: "Stephen Curry"
date: "10/29/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

#input
```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
head(wisc.df)
```

We have `r nrow(wisc.df)` samples in this dataset

```{r}
nrow(wisc.df)
```

#how many benign vs maligment
```{r}
table(wisc.df$diagnosis)
```


#convert features of data
```{r}
wisc.data <- as.matrix(wisc.df[,3:32])
row.names(wisc.data) <- wisc.df$id
head(wisc.data)
```

#I want to die
```{r}
diagnosis <- wisc.df$diagnosis
```

#Questions

Q1: 
```{r}
nrow(wisc.df)
```

Q2:
```{r}
table(wisc.df)
```

Q3
```{r}
colnames(wisc.df)
```

I can use `length()' to count how many matches the hoe has
```{r}
length(grep("_mean", colnames(wisc.df)))
```

```{r}
round(colMeans(wisc.data), 3)
```

```{r}
round(apply(wisc.data, 2, sd),3)
```

Them values look hella diff so im finna use `scale=TRUE` when I run that PCA

```{r}
#Do that PCA on wisc.data by yeeting the code
wisc.pr <- prcomp(wisc.data, scale=TRUE)
summary(wisc.pr)
```

```{r}
plot(wisc.pr)
```

Finna make a plot of PC1 vs. PC1=2

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2])
```

Color by cancer/mei you cancer

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2],, col=diagnosis)
```

Q4 #importance thing helps do stuff
```{r}
x <- summary(wisc.pr)
x$importance[,1]
```

Q5
```{r}
x$importance[3,] >0.7
```

```{r}
## ggplot based graph
library(factoextra)
fviz_eig(wisc.pr, addlabels = TRUE)
```

Q9:
```{r}

```

Q10:
```{r}
y <- summary(wisc.pr)
which(y$importance["Cumulative Proportion",] > 0.8)[1]
```


```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:3]))
```
```{r}
wisc.pr.hclust
```
```{r}
groups <- cutree(wisc.pr.hclust, k=2)
table(groups)
```

```{r}
grps <- cutree(wisc.pr.hlcust, k=2)
table(grps)
```

```{r}
table(groups, diagnosis)
```

```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
plot(wisc.pr$x[,1:2], col=diagnosis)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```

RED=cancer
Cancer=messed up cells cos science


