---
title: "Class08"
author: "Stephen Curry"
date: "10/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```


Use the kmeans() function setting k to 2 and nstart=20
```{r}
special_k <- kmeans(x, centers = 2, nstart = 20)
```

Inspect/print the results
```{r}
special_k
```

Q. How many points are in each cluster? #30
Q. What ‘component’ of your result object details
 - cluster size?
 - cluster assignment/membership? 
 - cluster center?
```{r}
special_k$size
```

```{r}
special_k$cluster
```
 
Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points

```{r}
plot(x, col = special_k$cluster)
points(special_k$centers, col = "turquoise", pch=15)
```
 

#Hierarchial clustering
hclust() requires distance matrix as input; get from dist() function

```{r}
# First we need to calculate point (dis)similarity
# as the Euclidean distance between observations
dist_matrix <- dist(x)
# The hclust() function returns a hierarchical
# clustering model
hc <- hclust(d = dist_matrix)
# the print method is not so useful here
hc 
```

```{r}
plot(hc)
abline(h=6, col="pink")
grps <- cutree(hc, h=6)
```
 
```{r}
cutree(hc, k=4)
```
 
```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)

```

 Q. Use the dist(), hclust(), plot() and cutree()
 functions to return 2 and 3 clusters
Q. How does this compare to your known 'col' groups?
```{r}
#clustering
hc <- hclust(dist(x))
#draw tree
plot(hc)
abline(h=2, col="pink")
#cut tree into clusters
groups<- cutree(hc, k=3)
groups
```
 
 

Plot data colored by their drip, 

How many points in each cluster
```{r}
table(groups)
```

Cross tabulate EX. compare clustering result with known answer; 

```{r}
table(groups, col)
```


#PCA stuffz 
Read example gene expression data; 
```{r}
mydata <- read.csv("https://tinyurl.com/expression-CSV",
 row.names=1)
head(mydata)
#how many genes in dataset
dim(mydata)
nrow(mydata)
```

```{r}
## lets do PCA with prcomp
pca <- prcomp(t(mydata), scale=TRUE) 

## See what is returned by the prcomp() function
attributes(pca)
```


```{r}
## A basic PC1 vs PC2 2-D plot
plot(pca$x[,1], pca$x[,2])
## Precent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1) 
```



```{r}
barplot(pca.var.per, main="YEET Plot",
 xlab="Principal Component", ylab="Percent Variation")
```


```{r}
plot(pca$x[,1], pca$x[,2])
  col=c("red", "red","red", "red", "red", "blue", "blue", "blue", "blue", "blue")
```


```{r}
x <- read.csv("UK_foods.csv", row.names = 1)
x
```


```{r}
barplot(as.matrix(x), beside=FALSE, col=rainbow(nrow(x)))
```

```{r}
pca <- prcomp(t(x))
summary(pca)
```

