---
title: "Class14"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eval=FALSE}
#eval=FALSE will cause chunk not to run?
install.packages("BiocManager")
BiocManager::install()
BiocManager::install("DESeq2")
```


```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```


```{r}
head(counts)
head(metadata)
```

How many genes in dataset?
```{r}
nrow(counts)
```

How many expts be?
```{r}
ncol(counts)
```

We want know if there be difference in expression values for control (non-drug) vs treated (i.e. drg added cell lines)

1st finna find which expts are control expts y get avg values across all control expts. Then do same for drug treated. 
```{r}
View(metadata)
```

```{r}
control$id
```

Calculate mean values across control columns of **countdata** This new robust code allows to do more than 4 groups
```{r}
control.mean2 <- rowSums(counts[,control$id])/length(control$id)
names(control.mean2) <- counts$ensgene

treated.mean2 <- rowSums(counts[,treated$id])/length(treated$id)
names(treated.mean2) <- counts$ensgene
```





```{r}
control <- metadata[metadata[,"dex"]=="control",]
control.mean <- rowSums( counts[ ,control$id] )/4 
names(control.mean) <- counts$ensgene

treated <- metadata[metadata[,"dex"]=="treated",]
treated.mean <- rowSums( counts[ ,treated$id] )/4 
names(treated.mean) <- counts$ensgene
```



```{r}
mean.counts2 <- data.frame(control.mean2, treated.mean2)
colSums(mean.counts2)
```


```{r}
#control.mean
mean.counts <- data.frame(control.mean, treated.mean)
colSums(mean.counts)
```

```{r}
plot(mean.counts2$control.mean2, mean.counts2$treated.mean2, log="xy")
```


```{r}
mean.counts2$log2fc <- 
  log2(mean.counts2[, "treated.mean2"]/mean.counts2[,"control.mean2"])
  head(mean.counts2)
```

```{r}
x <- c(56, 0, 26, 0)
x == 0
```

```{r}
which(x==0)
```

```{r}
x2 <- data.frame(control=c(56, 0), treated=c(26, 0))
x2
```






```{r}
x2==0
```

#arr.ind show exact location of where 0 is in matrix (row 2 col 1 )
```{r}
which(x2==0, arr.ind=TRUE)
```

```{r}
unique(which(x==0, arr.ind = ))
```

```{r}
zero.vals <- which(mean.counts2[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- mean.counts2[-to.rm,]
head(mycounts)
head(zero.vals)
```


```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
up.ind
```

```{r}
sum(up.ind)
```



```{r}
head(mycounts[up.ind,])
```


```{r}
library(DESeq2)
citation("DESeq2")
```

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds

dds <- DESeq(dds)
res <- results(dds)
res
```


```{r}
summary(res)
```

Volcano plot: combine Fold Change y p-value into one figure indication proportion of genes with large scale significant differences in expression
```{r}
#add color por favor
mycols <- rep("gray", nrow(res))
#make points +/-2 fold change blue
mycols[ abs(res$log2FoldChange) > 2 ] <- "blue"

#make points below p-value cutoff red
mycols[ (res$padj > 0.05)] <- "gray"

plot( res$log2FoldChange, -log(res$padj), col=mycols)
abline(v=c(-2,2), col="pink")
abline(h=-log(0.1), col="pink", lty=2)
```

Save our results
```{r}
write.csv(res, file="expression_results.csv")
```





































